---
title: "Determine redocking distance and grid score with increasing box size"
author: "Mehnaz Meem"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
    fig_width: 5
    fig_height: 5
theme: cosmo
---

Source: TBA

## Introduction

This analysis calculates the distance between the centre of the top-ranked
AutoDock Vina ligand pose and the crystal-ligand centre for 3ERT. It also
compares docking distance with Vina predicted binding affinity.

```{r,library}

library("beeswarm")

```

## Crystal structure position

```{r crystal}

cx <- 30.282
cy <- -1.913
cz <- 24.2065
cxyz <- c(cx, cy, cz)

```

## Read example data

```{r autodock_data}

pdb <- readLines("3ERT_docking.pdbqt", warn = FALSE)

# Keep only the top-ranked Vina pose (MODEL 1)
model_start <- grep("^MODEL[[:space:]]+1[[:space:]]*$", pdb)[1]

if (!is.na(model_start)) {
  model_end <- which(
    seq_along(pdb) > model_start & grepl("^ENDMDL", pdb)
  )[1]

  if (is.na(model_end)) {
    stop("MODEL 1 has no ENDMDL in the example PDBQT file")
  }

  pdb <- pdb[model_start:model_end]
}

atoms <- pdb[grep("^(ATOM|HETATM)", pdb)]

if (length(atoms) == 0) {
  stop("No ATOM or HETATM lines found in the example PDBQT file")
}

```

Extract the x, y and z coordinates from their fixed PDBQT positions.

```{r extract_coords}

x <- as.numeric(substr(atoms, 31, 38))
y <- as.numeric(substr(atoms, 39, 46))
z <- as.numeric(substr(atoms, 47, 54))

if (anyNA(c(x, y, z))) {
  stop("Could not read coordinates from the example PDBQT file")
}

mx <- cbind(x = x, y = y, z = z)

str(mx)
summary(mx)

```

## Predicted ligand centre and distance

```{r example_distance}

px <- mean(range(mx[, "x"]))
py <- mean(range(mx[, "y"]))
pz <- mean(range(mx[, "z"]))

dx <- abs(cx - px)
dy <- abs(cy - py)
dz <- abs(cz - pz)

hyp <- sqrt(dx^2 + dy^2 + dz^2)
hyp

```

## Define distance function

```{r dist_function}

model_ligand_dist <- function(cxyz, pdbqt_file) {
  pdb <- readLines(pdbqt_file, warn = FALSE)

  # Keep only the top-ranked Vina pose (MODEL 1)
  model_start <- grep("^MODEL[[:space:]]+1[[:space:]]*$", pdb)[1]

  if (!is.na(model_start)) {
    model_end <- which(
      seq_along(pdb) > model_start & grepl("^ENDMDL", pdb)
    )[1]

    if (is.na(model_end)) {
      stop(paste("MODEL 1 has no ENDMDL in:", pdbqt_file))
    }

    pdb <- pdb[model_start:model_end]
  }

  atoms <- pdb[grep("^(ATOM|HETATM)", pdb)]

  if (length(atoms) == 0) {
    stop(paste("No ATOM or HETATM lines found in:", pdbqt_file))
  }

  # Read coordinates from fixed-width PDBQT columns
  x <- as.numeric(substr(atoms, 31, 38))
  y <- as.numeric(substr(atoms, 39, 46))
  z <- as.numeric(substr(atoms, 47, 54))

  if (anyNA(c(x, y, z))) {
    stop(paste("Could not read coordinates from:", pdbqt_file))
  }

  predicted_centre <- c(
    mean(range(x)),
    mean(range(y)),
    mean(range(z))
  )

  sqrt(sum((cxyz - predicted_centre)^2))
}

```

Test the function.

```{r functest}

model_ligand_dist( cxyz, "3ERT_docking.pdbqt")

```

## Run for all PDBQT files

```{r runall}

myfiles <- list.files("3ERT_seed_run_docking",full.names=TRUE)

box_sizes <- sort(as.numeric(unique(sapply(strsplit(basename(myfiles),"_"),"[[",3))))
box_sizes

res <- lapply(box_sizes, function(x) {
  myfiles2 <- myfiles[grep(paste("_",x,"_",sep=""),myfiles)]
  res2 <- unlist(lapply(myfiles2,function(y) { model_ligand_dist(cxyz, y) } ) )
  return(res2)
})

names(res) <- box_sizes

boxplot(res, col="white", cex=0, xlab="Box size (A)", ylab="Redocking distance (A)",main="3ERT redocking")
beeswarm(res, pch=1, col="black", cex=0.2, add=TRUE)

lapply(res,summary)

```

## Extract Vina predicted binding affinity

```{r vina_energy}

extract_vina_energy <- function(pdbqt_file) {
  lines <- readLines(pdbqt_file, warn = FALSE)

  vina_line <- grep("^REMARK VINA RESULT:", lines, value = TRUE )

  if (length(vina_line) == 0) {
    return(NA_real_)
  }

  as.numeric( sub( "^REMARK VINA RESULT:[[:space:]]*([-0-9.]+).*", "\\1", vina_line[1] ) )
}

extract_vina_energy("3ERT_docking.pdbqt")

res0 <- lapply(box_sizes, function(x) {
  myfiles2 <- myfiles[grep(paste("_",x,"_",sep=""),myfiles)]
  res3 <- unlist(lapply(myfiles2,function(y) { extract_vina_energy(y) } ) )
  return(res3)
})

names(res0) <- box_sizes

boxplot(res0, col="white", cex=0, xlab="Box size (A)", ylab="Grid score",main="3ERT redocking")
beeswarm(res0, pch=1, col="black", cex=0.2, add=TRUE)

lapply(res0,summary)

```

## Negative control distance

```{r negdist}

myfiles <- list.files("3ERT_negative_docking",full.names=TRUE)

box_sizes <- sort(as.numeric(unique(sapply(strsplit(basename(myfiles),"_"),"[[",3))))
box_sizes

res4 <- lapply(box_sizes, function(x) {
  myfiles2 <- myfiles[grep(paste("_",x,"_",sep=""),myfiles)]
  res5 <- unlist(lapply(myfiles2,function(y) { model_ligand_dist(cxyz, y) } ) )
  return(res5)
})

names(res4) <- box_sizes

boxplot(res4, col="white", cex=0, xlab="Box size (A)", ylab="Redocking distance (A)",main="3ERT negative docking")
beeswarm(res4, pch=1, col="black", cex=0.2, add=TRUE)

lapply(res4,summary)

```

## Negative binding affinity

```{r neg_energy}

res6 <- lapply(box_sizes, function(x) {
  myfiles2 <- myfiles[grep(paste("_",x,"_",sep=""),myfiles)]
  res7 <- unlist(lapply(myfiles2,function(y) { extract_vina_energy(y) } ) )
  return(res7)
})

names(res6) <- box_sizes

boxplot(res6, col="white", cex=0, xlab="Box size (A)", ylab="Grid score",main="3ERT neg docking")
beeswarm(res6, pch=1, col="black", cex=0.2, add=TRUE)

lapply(res6,summary)

```



## Session information

```{r session}
sessionInfo()
```
